\documentclass[letterpaper, 11pt]{article}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{times}
\usepackage{listings}
\usepackage {tikz}
\usetikzlibrary {positioning}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{graphicx}
\usepackage[pdf]{graphviz}

\newtheorem{lemma}{Lemma}[section]
\newtheorem{figur}[lemma]{Figure}
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{proposition}[lemma]{Proposition}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{example}[lemma]{Example}
\newtheorem{exercise}[lemma]{Exercise}
\newtheorem{remark}[lemma]{Remark}
\newtheorem{fig}[lemma]{Figure}
\newtheorem{tab}[lemma]{Table}
\newtheorem{fact}[lemma]{Fact}
\newtheorem{test}{Lemma}
\newtheorem{algorithm}[lemma]{Algorithm}

\begin{document}



\title{Homework \#5 \\ Automata and Computation Theory \\Fall 2018}
\author{Written by Eric Rothman}

\maketitle

\section*{Problem 1}
This question studies the intersection and complementation of context free languages.
Consider the languages $A = \left\{ a ^ { m } b ^ { n } c ^ { n } | m , n \geq 0 \right\}$ and $B = \left\{ a ^ { n } b ^ { n } c ^ { m } | m , n \geq 0 \right\}$.
\subsection*{a}
Give a context-free grammar for each of $A$ and $B$. Then, use $A$ and $B$ to show that the class of context free languages is not closed under intersection.\\
$A$:\\

$S --> aS|C$

$C --> bCc|\epsilon$\\\\
$B$:\\

$S --> Sc|C$

$C --> aCb|\epsilon$\\\\
Lets look at $D = A \cap B$.
I will prove that $D = \{a^nb^nc^n | n \geq 0\}$
\subsubsection*{first direction}
$D \subset \{a^nb^nc^n | n \geq 0\}$

let $d \in D$.

So $d \in A$ and $d \in B$.

So $d$ is of the form $a ^ { m_1 } b ^ { n } c ^ { n }$ and of the form $a ^ { n } b ^ { n } c ^ { m_2 }$ where $n,m_1,m_2 \geq 0$.

So $m_1 = n$ since the number of $a$'s equals the number of $b$'s in the second form and $m_2 = n$ since the number of $b$'s is equal to the number of $c$'s in the first form.

So $m_1 = n = m_2$

So $d$ is of the form $a^nb^nc^n$, so $d \in \{a^nb^nc^n | n \geq 0\}$.
\subsubsection*{other direction}
$\{a^nb^nc^n | n \geq 0\} \subset D$

let $d \in \{a^nb^nc^n | n \geq 0\}$

So $d$ is of the form $a^nb^nc^n$.

So $d$ is of the form $a^mb^nc^n$ where $m = n$, so $d \in A$.

And $d$ is of the form $a^nb^nc^m$ where $m = n$, so $d \in B$.

Since $d \in A$ and $d \in B$, so $d \in A \cap B$, so $d \in D$.

So $D = \{a^nb^nc^n | n \geq 0\}$\\\\
We proved in class that $\{a^nb^nc^n | n \geq 0\}$ is a non context free language, but I'll rewrite it for prosperities sake.

\subsubsection*{Proof of nonCFL}
Assume $D$ is a CFL. Then $\exists$ a number $p$ s.t. the pumping lemma holds.
Let $S = a^pb^pc^p$.
Since $|S| = 3p > p$, the pumping lemma holds.

So $S$ can be divided into $S = uvxyz$.

$v$ can contain only one or two symbols since $|vxy| \leq p$ and every substring of length $p$ contains at most two symbols since each section of symbols is $p$ long.

$y$ can contain only one or two symbols for the same reason as $v$.

There are two cases we need to consider.\\

(1) Both $v$ and $y$ contain only one symbol.

Then atleast one symbol is missing in $vy$ and $|vy| > 0$
So $S' = uv^2xy^2z$ cannot have the same number of $a,b,c$'s
So there is a contradiction.\\

(2) Either $v$ or $y$ contain two symbols.

Since $|vxy| \leq p$ by the conditions of the pumping lemma, atleast one symbol is not in $vy$ since every substring of length $p$ contains at most two symbols since each section of symbols is $p$ long.

So $S' = uv^2xy^2z$ cannot have the same numbr of $a,b,c$'s
So there is a contradiction.\\

Since in either case there is a contradiction, the assumption must be false, so $D$ is not a context free language.

Since $A$ and $B$ are context free languages but $D = A \cap B$ is not a context free language, context free languages are not closed under intersection.

\subsection*{b}
Use (a) and DeMorgan’s Law (Textbook Theorem 0.20) to show that the class of context-free languages is not closed under complementation.

\subsubsection*{proof}
FSOC let the class of context free languages be closed under complement.

Let $A$ and $B$ be arbitrary context free languages.

Since CFL's are closed under complement and $A$ and $B$ are CFL's, $\bar{A}$ and $\bar{B}$ are CFL.

By what we proved in class, $C = \bar{A} \cup \bar{B}$ is also a context free language.

Since $C$ is a context free language, $\bar{C}$ is a context free language.

By DeMorgan's law, $\bar{C} =$ $!(\bar{A} \cup \bar{B}) = A \cap B$.

So $\bar{C}$ is equal to the intersection of two CFL's.

Since $A$ and $B$ are arbitrary context free languages and $\bar{C} = A \cap B$ is a CFL, the intersection of all context free languages is a CFl.

So CFL's are closed under intersection.\\

This contradicts what we proved in part a, so the assumption must be false.

So the class of CFL's are not closed under complement.
\newpage

\section*{Problem 2}
Let $D = \left\{ x y \left| x , y \in \{ 0,1 \} ^ { * } \text { and } \right| x | = | y | \text { but } x \neq y ^ { \mathcal { R } } \right\}$.
Give a context-free grammar for $D$, and formally prove that your grammar generates the give language (using the two directions argument).\\\\
$G$:

$S --> 0S0|1S1|A$

$A --> 0B1|1B0$

$B --> 0B0|1B1|0B1|1B0|\epsilon$

\subsection*{Proof}
We need to prove that $L(G) = D$, which we will do in the two directions.
\subsection*{First Direction}
$\forall w \in L(G)$, $w \in D$.\\

Observe that since every rule that adds a terminal adds two of them, one on either side of the current variable.

Also observe that there is only ever atmost one variable at a time during string generation and the generation must end with that variable becoming an $\epsilon$.

So $w$ can be divided into two sections, one for each side of the variable.

So $w = w_1w_2$ where the two substrings are divided by where the variable turned into an epsilon.

Since every time a terminal is added two terminals are added with one going into each substring, $|w_1| = |w_2|$.\\

Lets call $|w_1| = i$

Now observe that to stop generating the string, the variable must get the be $B$, the variable must be $A$ at some time.

During $A --> 0B1|1B0$, a different terminal is added to each substring mandatorily.

WLOG lets say that $A$ occurs at the $j$th spot in $w_1$, and the $i-j$ spot in $w_2$, because of how the transitions work the second substring is built backwards.

So $w_2^R$ contains the terminal added during the $A$ transition at the $i-(i-j) = j$th spot.

So the $j$th spot of $w_1$ and $w_2^R$ contains the terminal added by the $A$ transition.

Since $A$ adds a different terminal to each substring, the two characters at the $j$th spot of $w_1$ and $w_2^R$ are different charactors.

So $w_1 \neq w_2^R$, so $w = xy$ where $|x| = |y|$ and $x \neq y^R$.

So $w \in D$.
\subsection*{Other Direction}
prove $\forall w \in D$, $w \in L(G)$\\\\

since $w \in D$, $w = xy$ where $|x| = |y|$ and $x \neq y^R$.

Let $j$ be the first spot in the string $x$ where $x_j \neq y^R_j$.

This is guarenteed to occur since $x^R \neq y^R$, so there must be at least one spot where the charactors do not match.

So repeat $A --> 0A0|1A1$ until we arrive at the $j$th spot.
Since the grammer builds $y$ in reverse, every terminal added to the right is added as the next spot in $y^R$.

This matches $w$ since for every $i$th spot before the $j$th one, $x_i = y^R_i$.

At the $j$th spot transition to having the variable $A$ and then do the transition that matches $w$.

So if $x_j = 0$ do $A --> 0B1$ else if $x_j = 1$ do $A --> 1B0$.

Then the variable will be $B$, which can generate any strings $x$ and $y$ from that spot.

So just choose the transitions that build the $x$ and $y^R$ that you are looking for.

Then once the strings are build, do the transition to $\epsilon$ which ends the string generation.

So the string $w = xy$ was generated using $G$.

So $w \in L(G)$.

So $D \subset L(G)$.\\

Since $D \subset L(G)$ and $L(G) \subset D$, $D = L(G)$.

So $G$ is the correct grammer for generating $D$.
\newpage
\section*{Problem 3}
Prove that the following language over the alphabet $\Sigma = \{ 0,1,2 \}$ is not context free.
 
$C = \left\{ 0 ^ { a } 1 ^ { b } 2 ^ { c } | a , b , c \geq 0 \text { and } a > b \text { and } a > c \right\}$.\\
 
PROOF:

FSOC lets assume that $C$ is a CFL.
Then $\exists$ a number $p$ such that the pumping lemma holds.

Let $S = 0^{p+1}1^p2^p$.
$|S| = 3p+1 > p$ so the pumping lemma holds for this string.

So $S = uvxyz$

There are a two cases since both $v$ and $y$ can only contain up to two different symbols since $|vxy \leq p$ and each section of $S$ is of length atleast $p$.

\subsection*{Case 1}
$v$ or $y$ contains two symbols.

Then $S' = uv^2xy^2z$ is out of order since it would either contain the substring $0101$ or $1212$, which goes against the form that $C$ requires.

So $S' \notin C$ which contradicts the pumping lemma.

\subsection*{Case 2}
$v$ and $y$ contain only one symbol.

Within that there are three different cases.
\subsubsection*{Case a}
$v$ contains $0$.

If $v$ contains another symbol besides $0$, then it would fall within Case 1.

Since $v$ contains a $0$, $|v| = t > 0$.

Let $S' = uv^0xy^0z = uxz$.
Since $v$ is not in $S'$, the number of $0$'s in $S' = p-t \leq p$ since $t \geq 1$.

Now since $|vxy| < p$ and $b = p$, $y$ cannot contain any $2$'s without violating this condition.

So $S'$ contains as many $2$'s as $S$.
Let the number of $2$'s that $S'$ contain be called $c'$.

So $S'$ contains $p$ $2$'s.
Since the number of $0$'s in $S'$, named $a'$ WLOG, is less than or equal to $p$, and $c' = p$,
$a' \leq c'$.

So $S' \notin C$ which contradicts the pumping lemma.

\subsubsection*{Case b}
$v$ or $y$ contains $1$ and $v$ does not contain $0$.

let $t = |v| \geq 1$, or if $|v| = 0$, $t = |y| \geq 1$. 

Let $S' = uv^2xy^2z$

Let $b' =$ number of $1$'s in $S'$

Since either $v$ contains 1s or $y$ contains 1s, $b' = b+t > p$.

So $b' \geq p + 1$.

So $b' \geq a'$ since the number of $0$'s in $S'$ does not change.

So $S' \notin C$.

This contradicts the pumping lemma.
\subsubsection*{Case c}
$vy$ contain only $2$.

let $t = |vy| \geq 1$ by the conditions of the pumping lemma. 

Let $S' = uv^2xy^2z$

Let $c' =$ number of $2$'s in $S'$

Since either $vy$ contains $2$'s, $c' = p + t > p$

So $c' \geq p + 1$.

So $c' \geq a'$ since the number of $0$'s in $S'$ does not change.

So $S' \notin C$.

This contradicts the pumping lemma.\\

Since in every case there is a contradiction, the assumption must be false.

So $C$ is not a CFL.
\newpage
\section*{Problem 4}
Let $B$ be the language of all palindromes over $\{0, 1\}$ containing an equal number of $0$s and $1$s. Show that $B$ is not context-free.\\

PROOF:

FSOC lets assume that $C$ is a CFL.
Then $\exists$ a number $p$ such that the pumping lemma holds.

Let $S = 1^p0^{2p}1^p$
$|S| = 4p > p$, so the pumping lemma holds.

So $S = uvxyz$.

There are 3 cases.
\subsection*{Case 1}
$v$ or $y$ contain two symbols

It is impossible for both $v$ and $y$ to contain two symbols since $|vxy| \leq p$ and the distance between the two dividers, the only spot where $v$ or $y$ can contain two different symbols, is a distance of $2p$.

So WLOG let $v$ contain the two symbols.

Let $S' = uv^2xy^2z$.

Since $v$ contains two symbols, $S'$ is of the form $0^p1^t0^t1^{2p+k}0^p$ or $0^p1^{2p}0^t1^t0^{p+k}$ where $2t = |v|$ and $|y| = k$.

In either form, $S'$ is no longer a palidrome, so $S' \notin B$, which contradicts the pumping lemma.
\subsection*{Case 2}
$vy$ contain all of one symbol.

Then $v$ and $y$ are in the same segmant of $S$ since otherwise it would violate the fact that $vxy \leq p$ since there are $2p 0$'s.

Let $S' = uv^2xy^2z$.

If $vy$ just contains $0$'s, then there will be more $0$'s than $1$'s since $|vy| > 0$.

If $vy$ just contains $1$'s, then there will be more $1$'s than $0$'s since $|vy| > 0$.

Either way $S'$ does not contain an equal amount of $0$'s and $1$'s, so $S' \notin B$ which contradicts the pumping lemma.

\subsection*{Case 3}
$v$ and $y$ contain only one symbol each, but they contain different symbols.

Then either $v$ contains only $1$'s or $0$'s.

Let $S' = uv^2xy^2z$.

If $v$ contains only $1$'s, $S'$ has the form $0^p1^{2p+t}0^{p+j}$ where $t = |v|$ and $j = |y|$ where both are greater than zero otherwise it would be case 2.

If $v$ contains only $0$'s, $S'$ has the form $0^{p+t}1^{2p+j}0^p$ where $t = |v|$ and $j = |y|$ where both are greater than zero otherwise it would be case 2.

Either way, $S'$ is not longer a palindrome since its form reversed is different from when its normally read.

So $S' \notin B$ which contradicts the pumping lemma\\

Since in all cases there is a contradiction, the assumption must be false.

So $B$ is not a CFL.
\newpage
\end{document}